---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
options(tidyverse.quiet = TRUE)
```

# now

Experimental!

Forget about lifecycles, live in the {now}. 

`now::clean_up()` removes functions of given life cycles from the namespace's exports so they're not attached with `library()` and not available or suggested with `::`.

Why would we want that ?

* You might not want to use deprecated functions, to use best current practice
* You might not want to use experimental functions, because you need to write stable code
* You might not want to run into defunct functions, that exist only to fail
* You might want to explore a package without being distracted by outdated or unstable
  features.

In short:

* `clean_up(pkgs)` is `clean_up(pkgs, remove = getOption("now.lifecycles", c("superseded", "deprecated", "defunct")))`.
* Additional possible options are `"experimental"`, `"questioning"`, and `"reexports"`, 
  the latter is not a life cycle, but allows you to drop reexports so you're forced
  to use the source package when programming.
* For instance use`clean_up(pkg, lifecycle = "defunct")` to cleanup only defunct functions, and not break anything.
* Use`clean_up_tidyverse()` to do the above on all tidyverse packages at once.
* `clean_attach()` and `clean_attach_tidyverse()` are wrappers that call library

> **It's not meant to be used in production, because it will break some dependencies that rely on the
features you've removed**

It works by looking at the doc and the code of a function, but we might miss
some "life-cycled" functions if the doc doesn't clearly use life cycle tags and the code
doesn't call a {lifecycle} function directly.

## Installation

Install with

``` r
remotes::install_github("moodymudskipper/now")
```

## Example

```{r, error = TRUE}
library(tidyverse)
before <- sapply(tidyverse_packages(), \(pkg) length(getNamespaceExports(pkg)))
do
location

now::clean_up_tidyverse()
after <- sapply(tidyverse_packages(), \(pkg) length(getNamespaceExports(pkg)))
do
location
dplyr::do

# expect some failures! Many packages use superseded functions in particular
library(dm)
```

To avoid the latter, either :

* call `library(dm)` before cleaning up (these dependencies would
be copied in dm's imports env so `dm::separate()` would work but `tidyr::separate()` wouldn't)
* call `clean_up_tidyverse(force_keep = c("separate", "transmute"))` above to keep the necessary dependencies around.
* call `clean_up("dm")` or `clean_attach("dm")` before `clean_up_tidyverse()` and
don't be bothered with {dm}'s re-exports

Let's take a look at how much we can expect to trim off

```{r, error = TRUE}
# removing only superseded/deprecated/defunct
left_join(enframe(before, value = "before"), enframe(after, value = "after"), by = "name") |> 
  filter(after != before) |>
  mutate(keep_ratio = sprintf("%2.0f%%", 100*after/before)) |>
  arrange(keep_ratio) |>
  knitr::kable()

# removing additionally all experimental and questioning features, along with reexports
now::clean_up_tidyverse(c("experimental", "superseded", "deprecated", "defunct", "questioning", "reexports"))
after <- sapply(tidyverse_packages(), \(pkg) length(getNamespaceExports(pkg)))
left_join(enframe(before, value = "before"), enframe(after, value = "after"), by = "name") |> 
  filter(after != before) |>
  mutate(keep_ratio = sprintf("%2.0f%%", 100*after/before)) |>
  arrange(keep_ratio) |>
  knitr::kable()
```

